---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv(unzip("activity.zip"))
data$date <- as.Date(data$date)

```


## What is mean total number of steps taken per day?
```{r}
steps_per_day <- aggregate(x = data$steps, by = list(data$date), FUN=sum)
hist(steps_per_day$x, main="Histogram of steps per day", xlab="steps")
```

Mean steps per day:
```{r}
mean(steps_per_day$x, na.rm = TRUE)
```

Median steps per day:
```{r}
median(steps_per_day$x, na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r}
mean_activity <- aggregate(x = data$steps, by = list(data$interval), FUN=function(x) mean(x, na.rm = TRUE))
with(mean_activity, plot(x ~ Group.1, type='l'))
```

Which 5-minute interval has the highest average number of steps?
```{r}
mean_activity[which.max(mean_activity$x),'Group.1']
```

## Imputing missing values
How many NA's are there in the dataset?
```{r}
sum(is.na(data))
```

Replace all missing values with mean step value from all days for that 5-minute interval
```{r}
library(dplyr)
impute_mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

nona_data <- data %>% group_by(interval) %>% mutate(steps = impute_mean(steps))
```
Perform the histogram, mean and median analyses with the new dataset
```{r}
steps_per_day2 <- aggregate(x = nona_data$steps, by = list(nona_data$date), FUN=sum)
hist(steps_per_day2$x, main="Histogram of steps per day, NA's imputed", xlab="steps")
```

```{r}
mean(steps_per_day2$x)
```

```{r}
median(steps_per_day2$x)
```
## Are there differences in activity patterns between weekdays and weekends?

Create a new variable to mark whether a date is on the weekend or not
```{r}
weekend=c("Saturday", "Sunday")
nona_data <- nona_data %>% mutate (is_weekend = as.factor(weekdays(date) %in% weekend))
levels(nona_data$is_weekend) <- c("weekday", "weekend")
plotdata <- nona_data %>% group_by(is_weekend, interval) %>%summarize(mean_step = mean(steps))

library(ggplot2)
qplot(x=interval, y=mean_step, facets = is_weekend ~ ., data=plotdata, geom="line")
```
